---
title: "HW5 - Make Data Move"
author: ""
date: "05/10/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#  Explore the recent global developments with R
Today, you will load a filtered gapminder dataset -  with a subset of data on global development from 1952 - 2007 in increments of 5 years - to capture the period between the Second World War and the Global Financial Crisis. 

**Your task: Explore the data and visualise it in both static and animated ways, providing answers and solutions to 7 questions/tasks below.**

## Get the necessary packages
First, start with installing the relevant packages 'tidyverse', 'gganimate', and 'gapminder'.

```{r libraries, echo = FALSE}
library(tidyverse)
library(gganimate)
library(gapminder)
```

## Look at the data
First, see which specific years are actually represented in the dataset and what variables are being recorded for each country. Note that when you run the cell below, Rmarkdown will give you two results - one for each line - that you can flip between.
```{r}
unique(gapminder$year)
head(gapminder)
```
The dataset contains information on each country in the sampled year, its continent, life expectancy, population, and GDP per capita.

Let's plot all the countries in 1952.
```{r 1957}
theme_set(theme_bw())  # set theme to white background for better visibility

ggplot(subset(gapminder, year == 1952), aes(gdpPercap, lifeExp, size = pop)) +
  geom_text(aes(label = country)) +
  scale_x_log10()
```
We see an interesting spread with an outlier to the right. Answer the following questions, please:

Q1. Why does it make sense to have a log10 scale on x axis?

-  The outlier in the data makes it necessary to use log10 in order for the rest of the visualisation to still be meaningful and detailed enough

Q2. What country is the richest in 1952 (far right on x axis)? 

- I use the geom_text function to find the richest country in 1952:geom_text(aes(label = country)). In this way the country name is printet out instead of showing a dot.The answer is Kuwait as it is the country far right on the x axis. 

You can generate a similar plot for 2007 and compare the differences
```{r 2007}
ggplot(subset(gapminder, year == 2007), aes(gdpPercap, lifeExp, size = pop)) +
  geom_point(aes(color = continent)) +
  scale_x_log10() +
  labs(titel = "Country Wealth", x = "life expectancy at birth", y = "per-capita GDP")
```
The black bubbles are a bit hard to read, the comparison would be easier with a bit more visual differentiation.

Q3. Can you differentiate the continents by color and fix the axis labels?

- I use geom_point(aes(color = continent) to differentiate the continents by color. I use labs(titel = "Country Wealth", x = "life expectancy at birth", y = "per-capita GDP") to change the axis labels.    

Q4. What are the five richest countries in the world in 2007?

- I first use the gapminder pipe: gapminder %>%, then I filter by years so the year must be 2007: filter(year == 2007) %>%, then I arrange the countries by descenting order according to the gdpPercap value arrange(desc(gdpPercap)).    


```{r}
gapminder %>%
filter(year == 2007) %>% 
arrange(desc(gdpPercap))
```

- The five richest countries in the world in 2007 is Norway, Kuwait, Singapore, United States and Ireland.
##  Make it move!

The comparison would be easier if we had the two graphs together, animated. We have a lovely tool in R to do this: the `gganimate` package. And there are two ways of animating the gapminder ggplot.

### Option 1: Animate using transition_states() 
install.packages('av')
library(av)
The first step is to create the object-to-be-animated
```{r anim1 start}
anim <- ggplot(gapminder, aes(gdpPercap, lifeExp, size = pop)) +
  geom_point() +
  scale_x_log10()  # convert x to log scale
anim
```

This plot collates all the points across time. The next step is to split it into years and animate it. This may take some time, depending on the processing power of your computer (and other things you are asking it to do). Beware that the animation might appear in the 'Viewer' pane, not in this rmd preview. You need to knit the document to get the viz inside an html file.
```{r anim1}
anim + transition_states(year, 
                      transition_length = 1,
                      state_length = 1)

```
Notice how the animation moves jerkily, 'jumping' from one year to the next 12 times in total. This is a bit clunky, which is why it's good we have another option. 


### Option 2 Animate using transition_time()
This option smoothes the transition between different 'frames', because it interpolates and adds transitional years where there are gaps in the timeseries data.

```{r anim2}
options(pillar.sigfigs=100)
anim2 <- ggplot(gapminder, aes(gdpPercap, lifeExp, size = pop)) +
  geom_point() +
  transition_time(year)+
  labs(title = "Year: {frame_time}", x = "life expectancy at birth", y = "per-capita GDP" )
anim2
```

The much smoother movement in Option 2 will be much more noticeable if you add a title to the chart, that will page through the years corresponding to each frame.


Q5 Can you add a title to one or both of the animations above that will change 
 in sync with the animation? [hint: search labeling for transition_states() and transition_time() functions respectively]
 
- I added the function labs(title = "Year: {frame_time}") to the last animation, in this way the years will change in sync with the animation. 

Q6 Can you make the axes' labels and units more readable? Consider expanding the abreviated lables as well as the scientific notation in the legend and x axis to whole numbers.[hint:search disabling scientific notation]

- I change the names of the x and y axis to life expectancy at birth and per-capita GDP. labs(title = "Year: {frame_time}", x = "life expectancy at birth", y = "per-capita GDP"). I change the scientific notation to whole numbers by adding options(pillar.sigfigs=100) to the code. By doing this I change the number of significant digits that will be printed.  

Q7 Come up with a question you want to answer using the gapminder data and write it down. Then, create a data visualisation that answers the question and explain how your visualization answers the question. (Example: you wish to see what was mean life expectancy across the continents in the year you were born versus your parents' birth years). [hint: if you wish to have more data than is in the filtered gapminder, you can load either the `gapminder_unfiltered` dataset and download more at https://www.gapminder.org/data/ ]

- Has life expectancy in Europe increased from 1952 to 2007?
- I have made a data visualization that shows the years 1952 to 2007 and the life expectancy in Europe. Here one can see that life expectancy has generally increased over the years. This can be seen by looking at the points that are lowest in value on the x axis. Here it is seen that the lowest expected age increases. It is especially in the years from 1952 to 1980 that the increase takes place. After this, it seems that there is a slight decline in the lowest life expectancy, but at the same time it is seen that the highest life expectancy is increasing. This means that there is a greater spread between the lowest and highest life expectancy in Europe.

```{r Europe}
ggplot(subset(gapminder, continent == 'Europe'), aes(lifeExp, year)) +
  geom_point(aes(color = country))+
  labs(titel = 'life expectancy at birth in Europe', x = 'life expectancy at birth')
```

    
    
    
