title: "W8 practicing web scraping"
author: "Johanne Würtz"
date: "11/2/2020"
output: html_document

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r libraries}
library(rvest)
library(dplyr)
library(tidyr)
library(stringr)
library(janitor)
```

I want to do task 2 of the week's tasks. I could not think of an interest, but I came across a website with the 500 most popular dog names and thought why not?

The website can be found with the following link: https://www.puppyleaks.com/popular-dog-names/

```{r }
url <- 'https://www.puppyleaks.com/popular-dog-names/'
url_html <- read_html(url)
```

I'm new to web scraping, so I reused Adela's code to get a better sense of what the different chunks of code can do.

```{r extract the individual rows}
whole_table <- url_html %>% 
 html_nodes("tr") %>%
 html_text(trim = FALSE) 
head(whole_table)
```

```{r extract individual cells}
whole_table <- url_html %>% 
 html_nodes("td") %>% #en individuel celle (td)
 html_text(trim = FALSE) 
head(whole_table)
```

```{r extract the whole HTML table}
whole_table <- url_html %>% 
 html_nodes("table") %>%
 html_table(fill = TRUE)  #str(whole_table) turns out to be a list
```

By using head() and tail() I get to look at the first and last 6 rows in the table 

```{r}
new_table <- do.call(cbind,unlist(whole_table, recursive = FALSE)) 
head(new_table)
tail(new_table)
```

I create a csv file with the name PopularDogNames.csv to make it easier to use the data later.

```{r write csv}
write.csv(new_table,"PopularDogNames.csv")
```

To check out the columns in the table

```{r}
colnames(new_table)
```

The table contains information about female dog names and male dog names and is sorted alphabetically.









